---
title: 'STAT 231: Problem Set 2B'
author: "Matthew Perkins"
date: "due by 5 PM on Friday, March 5"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

Series B homework assignments are designed to help you further ingest and practice the material covered in class over the past week(s).  You are encouraged to work with other students, but all code must be written by you and you must indicate below who you discussed the assignment with (if anyone).  

Steps to proceed:

\begin{enumerate}
\item In RStudio, go to File > Open Project, navigate to the folder with the course-content repo, select the course-content project (course-content.Rproj), and click "Open" 
\item Pull the course-content repo (e.g. using the blue-ish down arrow in the Git tab in upper right window)
\item Copy ps2B.Rmd from the course repo to your repo (see page 6 of the GitHub Classroom Guide for Stat231 if needed)
\item Close the course-content repo project in RStudio
\item Open YOUR repo project in RStudio
\item In the ps2B.Rmd file in YOUR repo, replace "YOUR NAME HERE" with your name
\item Add in your responses, committing and pushing to YOUR repo in appropriate places along the way
\item Run "Knit PDF" 
\item Upload the pdf to Gradescope.  Don't forget to select which of your pages are associated with each problem.  \textit{You will not get credit for work on unassigned pages (e.g., if you only selected the first page but your solution spans two pages, you would lose points for any part on the second page that the grader can't see).} 
\end{enumerate}

```{r, setup, include=FALSE}
library(mdsr)   
library(tidyverse)
library(Lahman)

knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```


\newpage 

# If you discussed this assignment with any of your peers, please list who here:

> ANSWER:

\newpage
# MDSR Exercise 4.14 (modified)

Use the `Pitching` data frame from the `Lahman` package to identify every pitcher in baseball history who has accumulated at least 300 wins (`W`) and at least 3,000 strikeouts (`SO`).  

a. How many pitchers meet this criteria?  

> ANSWER: 10 pitchers meet this criteria.

```{r}
PitchingTotals <- Pitching %>%
  select(playerID,stint,W,SO) %>%
  group_by(playerID) %>%
  summarize(
    total_W=sum(W),
    total_SO=sum(SO))%>%
    arrange(desc(total_SO))%>%
  filter(total_W>299 & total_SO > 3000)
PitchingTotals
```

b. Which of these pitchers had the most accumulated strikeouts?  How many strikeouts had he accumulated?  What is the most strikeouts he had in one season?

> ANSWER: Nolan Ryan accumulated the most strikeouts, with a total of 5714. In the 1973 season, he had 383 strikeouts.

```{r}
Nol <- Pitching %>%
  filter(playerID == "ryanno01") %>%
    arrange(desc(SO))
Nol
```


\newpage 
# MDSR Exercise 4.17 (modified)

a. The Violations data set in the `mdsr` package contains information regarding the outcome of health inspections in New York City.  Use these data to calculate the median violation score by zipcode and dba for zipcodes in Manhattan.  What pattern (if any) do you see between the number of inspections and the median score?  Generate a visualization to support your response.  

> ANSWER: While the pattern is not totally clear, it appears as though the median violation score by zipcode or by dba tends to increase as inspections increase.

```{r}
ViolationsZip <- Violations %>%
  select(zipcode,score) %>%
  filter(is.na(score)==FALSE)%>%
  group_by(zipcode) %>%
  summarize(
    total_insp = length(zipcode),
    med_score = median(score)) %>%
    arrange(desc(med_score))
ViolationsZip

ggplot(data = ViolationsZip) +
    geom_point(mapping = aes(
      x = total_insp,
      y = med_score))+
      geom_smooth(aes(x = total_insp, 
      y = med_score), method = "lm")
      

ViolationsDba <- Violations %>%
  select(dba,score,zipcode) %>%
  filter(is.na(score)==FALSE)%>%
  group_by(dba,zipcode) %>%
  summarize(
    total_insp = length(dba),
    med_score=median(score))%>%
    arrange(desc(med_score))
ViolationsDba

ggplot(data = ViolationsDba) +
    geom_point(mapping = aes(
      x = total_insp,
      y = med_score))+
      geom_smooth(aes(x = total_insp, 
      y = med_score), method = "lm")
```

b.  In your visualization in part (a), there should be at least a few points that stand out as outliers.  For *one of the outliers*, add text to the outlier identifying what business it is and an arrow pointing from the text to the observation.  First, you may want to `filter` to identify the name of the business (so you know what text to add to the plot).  

(Can't remember how to create a curved arrow in `ggplot`?  The answers to [this question](https://stackoverflow.com/questions/38008863/how-to-draw-a-nice-arrow-in-ggplot2/61383034) on Stack Exchange may help.  Can't remember how to add text to the plot in `ggplot`?  Check out the text examples with `annotate` [here](https://ggplot2.tidyverse.org/reference/annotate.html), or answers to [this question](https://stackoverflow.com/questions/14351608/color-one-point-and-add-an-annotation-in-ggplot2/14351810) that use `geom_text`.)

```{r}
filter(ViolationsDba, total_insp>200)
ggplot(data = ViolationsDba) +
    geom_point(mapping = aes(
      x = total_insp,y=med_score))+
      geom_smooth(aes(x = total_insp, 
      y = med_score), method = "lm")+
  geom_curve(aes(x=200,y=35,xend = 218,yend=15)
             ,arrow = arrow(length = unit(0.03,
              "npc")))+
  geom_text(x=190,y=40,label = 'Crown Fried
            Chicken',size=3)
```


\newpage
# MDSR Exercise 5.7

Generate the code to convert the data frame shown with this problem in the textbook (on page 130, and shown below) to wide format (i.e., the result table).  Hint: use `gather()` in conjunction with `spread()`; OR `pivot_longer()` in conjunction with `pivot_wider()`.

```{r}
#Didn't use pivotlonger, but this got the job done!

FakeDataLong <- data.frame(grp = c("A","A","B", "B")
                           , sex = c("F", "M", "F", "M")
                           , meanL = c(0.22, 0.47, 0.33, 0.55)
                           , sdL = c(0.11, 0.33, 0.11, 0.31)
                           , meanR = c(0.34, 0.57, 0.40, 0.65)
                           , sdR = c(0.08, 0.33, 0.07, 0.27))

DataWide <- FakeDataLong %>%
  pivot_wider(
    names_from = sex, 
    values_from = c(meanL,meanR,sdL,meanR,sdR), 
    values_fill = 0) %>%
    select(grp, F.meanL = meanL_F, 
           F.meanR = meanR_F,
           F.sdL = sdL_F, 
           F.sdR = sdR_F, 
           M.meanL = meanL_M, 
           M.meanR = meanR_M, 
           M.sdL = sdL_M, M.sdR = sdR_M)
DataWide
```

\newpage
# PUG Brainstorming

What topics or questions are you interested in exploring related to your PUG theme?  Dream big here.  Don't worry about whether there is data out there that's available and accessible that you could use to address your questions/topics.  Just brainstorm some ideas that get you excited.  Then, email your PUG team with your ideas.  Title the email "PS2B Brainstorming: PUG [#] [Topic]" and CC me (kcorreia@amherst.edu) on the email.  If another PUG member already initiated the email, reply all to their email.  

If you don't remember your PUG # and Topic, please see the file "PUGs" on the Moodle page under this week.  

If you don't know your PUG members email address, go to the class's Google group conversations (e.g., by clicking the link "Link to Google group conversations" at the top of our Moodle course page).  Then, on the navigation panel (left hand side), select "Members". 

> ANSWER: Do not write anything here.  Email your ideas to your PUG team and me in a message titled "PS2B Brainstorming: PUG [#] [Topic]".
