---
title: "STAT 231: Problem Set 6B"
author: "Matthew Perkins"
date: "due by 10 PM on Friday, April 2"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

This homework assignment is designed to help you further ingest, practice, and expand upon the material covered in class over the past week(s).  You are encouraged to work with other students, but all code and text must be written by you, and you must indicate below who you discussed the assignment with (if anyone).  

Steps to proceed:

\begin{enumerate}
\item In RStudio, go to File > Open Project, navigate to the folder with the course-content repo, select the course-content project (course-content.Rproj), and click "Open" 
\item Pull the course-content repo (e.g. using the blue-ish down arrow in the Git tab in upper right window)
\item Copy ps6B.Rmd from the course repo to your repo (see page 6 of the GitHub Classroom Guide for Stat231 if needed)
\item Close the course-content repo project in RStudio
\item Open YOUR repo project in RStudio
\item In the ps6B.Rmd file in YOUR repo, replace "YOUR NAME HERE" with your name
\item Add in your responses, committing and pushing to YOUR repo in appropriate places along the way
\item Run "Knit PDF" 
\item Upload the pdf to Gradescope.  Don't forget to select which of your pages are associated with each problem.  \textit{You will not get credit for work on unassigned pages (e.g., if you only selected the first page but your solution spans two pages, you would lose points for any part on the second page that the grader can't see).} 
\end{enumerate}

```{r, setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(wordcloud)
library(textdata)
library(janitor)

knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```


\newpage 
# If you discussed this assignment with any of your peers, please list who here:

> ANSWER:

\newpage
# Trump Tweets

David Robinson, Chief Data Scientist at DataCamp, wrote a blog post ["Text analysis of Trump's tweets confirms he writes only the (angrier) Android half"](http://varianceexplained.org/r/trump-tweets/).

He provides a dataset with over 1,500 tweets from the account realDonaldTrump between 12/14/2015 and 8/8/2016.  We'll use this dataset to explore the tweeting behavior of realDonaldTrump during this time period.

First, read in the file. Note that there is a `TwitteR` package which provides an interface to the Twitter web API.  We'll use this R dataset David created using that package so that you don't have to set up Twitter authentication.  

```{r}
load(url("http://varianceexplained.org/files/trump_tweets_df.rda"))
```

## A little wrangling to warm-up

1a.  There are a number of variables in the dataset we won't need.  

- First, confirm that all the observations in the dataset are from the screen-name `realDonaldTrump`.  

- Then, create a new dataset called `tweets` that only includes the following variables:

- `text`
- `created`
- `statusSource`

```{r}
trump_tweets_df
tweets_test <- trump_tweets_df %>%
  filter(screenName == 'realDonaldTrump') 
tweets_test #same number of rows, so must be all RDT tweets

tweets <- trump_tweets_df %>%
  select(text, created, statusSource)
tweets

```

\newpage
1b. Using the `statusSource` variable, compute the number of tweets from each source.  How many different sources are there?  How often are each used?

> ANSWER: There are 5 different sources, with android used for 762 tweets,
iphone used for 628 tweets, the web client used for 120 tweets, and 1 tweet
each via ipad and instagram.

```{r}
tweetSource <- tweets %>%
  pivot_wider(names_from = statusSource, values_from = text) %>%
  clean_names() %>%
  rename(android = a_href_http_twitter_com_download_android_rel_nofollow_twitter_for_android_a,
         iphone = a_href_http_twitter_com_download_iphone_rel_nofollow_twitter_for_i_phone_a, 
         web = a_href_http_twitter_com_rel_nofollow_twitter_web_client_a,
         ipad = a_href_http_twitter_com_number_download_ipad_rel_nofollow_twitter_for_i_pad_a,
         insta = a_href_http_instagram_com_rel_nofollow_instagram_a) %>%
  summarise(android_tweets = sum(!is.na(android)),
            iphone_tweets = sum(!is.na(iphone)),
            web_tweets = sum(!is.na(web)),
            ipad_tweets = sum(!is.na(ipad)),
            insta_tweets = sum(!is.na(insta)))
tweetSource
```

\newpage
1c. We're going to compare the language used between the Android and iPhone sources, so only want to keep tweets coming from those sources.  Explain what the `extract` function (from the `tidyverse` package) is doing below.  Include in your own words what each argument is doing.  (Note that "regex" stands for "regular expression".)

> ANSWER: The extract function is isolating which platform a tweet is from based
on the information contained within the statusSource column. First, col = 
statusSource specifies the statusSource column. into = "source" specifies the
name of the new column to be created. Regex is used to extract the desired 
sourceinformation, capturing charachers after twitter for and before the <. 
Remove = FALSE tells R to leave the original column in the dataframe.

```{r}
tweets2 <- tweets %>%
  extract(col = statusSource, into = "source"
          , regex = "Twitter for (.*)<"
          , remove = FALSE) %>%
  filter(source %in% c("Android", "iPhone"))
tweets2
```


\newpage
## How does the language of the tweets differ by source?  

2a. Create a word cloud for the top 50 words used in tweets sent from the Android.  Create a second word cloud for the top 50 words used in tweets sent from the iPhone.  How do these word clouds compare?  (Are there some common words frequently used from both sources? Are the most common words different between the sources?)

*Don't forget to remove stop words before creating the word cloud.  Also remove the terms "https" and "t.co".*

> ANSWER: The tweets from the android and the iphone have many distinct 
properties. While both share many words in common such as hillary and maga, the
frequencies are quite different. For the android tweets which Trump presumably 
sends himself, Hillary is at the top of the most common words along with other
political rivals, whereas the iPhone tweets appear to be more focused toward 
promoting the campaign with Trump2016 and maga as the clear frontrunners.

```{r, fig.width=8, fig.height=8}
tweets_words <- tweets2 %>%
  unnest_tokens(output = word, input = text)

tweets_words2 <- tweets_words %>%
  filter(!word %in% stop_words$word) %>%
  filter(word != "t.co", word != "https" & source == "Android")

word_frequencies <- tweets_words2 %>%
  count(word, sort = TRUE)
  
word_frequencies %>%
  with(wordcloud(words = word, freq = n, max.words=50))

tweets_words3 <- tweets_words %>%
  filter(!word %in% stop_words$word) %>%
  filter(word != "t.co", word != "https" & source == "iPhone")

word_frequencies2 <- tweets_words3 %>%
  count(word, sort = TRUE)
  
word_frequencies2 %>%
  with(wordcloud(words = word, freq = n, max.words=50))


```

\newpage
2b. Create a visualization that compares the top 10 *bigrams* appearing in tweets by each source (that is, facet by source).  After creating a dataset with one row per bigram, you should remove any rows that contain a stop word within the bigram.  

How do the top used bigrams compare between the two sources?

> ANSWER: There is a fair amount of overlap among bigrams (e.g. crooked hillary
and hillary clinton are near the top on both), but the iPhone tweets show a much
stronger focus on slogans such as makeamericagreatagain and trump2016.

```{r}
data(stop_words)

tweets2

tweets_bigrams_android <- tweets2 %>%
  filter(source == "Android") %>%
  unnest_tokens(output = bigram, input = text
                , token = "ngrams", n = 2) %>%
  count(bigram, sort = TRUE) %>%
  slice(1:100)

my_regex <- regex(paste("\\b", stop_words$word, "\\b", sep = "", collapse = "|"))

tweets_android_filtered <- tweets_bigrams_android %>%
  filter(!str_detect(bigram, my_regex), !str_detect(bigram, "https")) %>%
  slice(1:10)

tweets_android_filtered

tweets_bigrams_iphone <- tweets2 %>%
  filter(source == "iPhone") %>%
  unnest_tokens(output = bigram, input = text
                , token = "ngrams", n = 2) %>%
  count(bigram, sort = TRUE) %>%
  slice(1:100)

tweets_iphone_filtered <- tweets_bigrams_iphone %>%
  filter(!str_detect(bigram, my_regex), !str_detect(bigram, "https")) %>%
  slice(1:10)

tweets_iphone_filtered


tweets_android_filtered %>%
  ggplot(aes(x = reorder(bigram,n), y = n, color = bigram, fill=bigram)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  labs(y = "Number of instances"
       , title="The most common bigrams in Trump's Android tweets") +
  guides(color = "none", fill = "none")

tweets_iphone_filtered %>%
  ggplot(aes(x = reorder(bigram,n), y = n, color = bigram, fill=bigram)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  labs(y = "Number of instances"
       , title="The most common bigrams in Trump's iPhone tweets") +
  guides(color = "none", fill = "none")




```


\newpage
2c. Consider the sentiment.  Compute the proportion of words among the tweets within each source classified as "angry" and the proportion of words classified as "joy"  based on the NRC lexicon.  How does the proportion of "angry" and "joy" words compare between the two sources?  What about "positive" and "negative" words?  

> ANSWER: The proportion of anger and negative words according to nrc word sentiments is higher among android
tweets than iphone tweets. Anger words make up ~2.3% of android tweets and only 1.7% of iphone. Joy values are
roughly the same in both. Likewise, Negative words are ~4.2% of android tweets and only 2.7% of iphone. There
are also slightly fewer positive words in android tweets, but this difference is fairly small.

```{r}
nrc_lexicon <- get_sentiments("nrc")

nrc_angry <- nrc_lexicon %>%
  filter(sentiment == "anger")

nrc_joy <- nrc_lexicon %>%
  filter(sentiment == "joy")

tweets_words_android <- tweets2 %>%
  unnest_tokens(output = word, input = text) %>%
  filter(source == "Android")

tweets_words_android
total_android = 15764

tweets_words_android_angry <- tweets_words_android %>%
  inner_join(nrc_angry) %>%
  count(word, sort = TRUE, name = 'angry')
  
tweets_words_android_angry %>%
  summarize(sum(angry)/total_android)

tweets_words_android_joy <- tweets_words_android %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE, name = 'joy')

tweets_words_android_joy %>%
  summarize(sum(joy)/total_android)

tweets_words_iphone <- tweets2 %>%
  unnest_tokens(output = word, input = text) %>%
  filter(source == "iPhone")

tweets_words_iphone
total_iphone = 9632

tweets_words_iphone_angry <- tweets_words_iphone %>%
  inner_join(nrc_angry) %>%
  count(word, sort = TRUE, name = 'angry')
  
tweets_words_iphone_angry %>%
  summarize(sum(angry)/total_iphone)

tweets_words_iphone_joy <- tweets_words_iphone %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE, name = 'joy')

tweets_words_iphone_joy %>%
  summarize(sum(joy)/total_iphone)






nrc_negative <- nrc_lexicon %>%
  filter(sentiment == "negative")

nrc_positive <- nrc_lexicon %>%
  filter(sentiment == "positive")


tweets_words_android_negative <- tweets_words_android %>%
  inner_join(nrc_negative) %>%
  count(word, sort = TRUE, name = 'negative')
  
tweets_words_android_negative %>%
  summarize(sum(negative)/total_android)

tweets_words_android_positive <- tweets_words_android %>%
  inner_join(nrc_positive) %>%
  count(word, sort = TRUE, name = 'positive')

tweets_words_android_positive %>%
  summarize(sum(positive)/total_android)

tweets_words_iphone_negative <- tweets_words_iphone %>%
  inner_join(nrc_negative) %>%
  count(word, sort = TRUE, name = 'negative')
  
tweets_words_iphone_negative %>%
  summarize(sum(negative)/total_iphone)

tweets_words_iphone_positive <- tweets_words_iphone %>%
  inner_join(nrc_positive) %>%
  count(word, sort = TRUE, name = 'positive')

tweets_words_iphone_positive %>%
  summarize(sum(positive)/total_iphone)




```


\newpage
2d. Lastly, based on your responses above, do you think there is evidence to support Robinson's claim that Trump only writes the (angrier) Android half of the tweets from realDonaldTrump?  In 2-4 sentences, please explain.

> ANSWER: There does appear to be evidence to support Robinson's claim that Trump only writes the android half
of tweets from realDonaldTrump. At the very least, we can identify a clear difference in sentiments and preferred
bigrams between the two sources which is a strong indicator that they are from different authors. Especially given
the focus of iphone tweets on campaign slogans as opposed to phtases such as "lyin ted", it does seem likely that
Trump is in charge of the angrier tweets while a social media individual is tweeting the others.

